### 2013-12-18 paper/main.tex
moved Control Response section to start of Appendix
documented Control Law intuitive design
The Control Law intuitive design has been fully documented
with block diagrams and plots.

\usepackage{tikz}
\usetikzlibrary{calc,arrows,positioning,shadows}
%
% control system tikz objects
\tikzset{shadow/.style={drop shadow,fill=white}}
\tikzset{sum/.style={circle,draw,very thick,shadow,minimum size=5mm}}
\tikzset{sumnp/.style={sum,
				label={{above left,xshift=0.0mm,yshift=-1.0mm}:$+$},
				label={{below right,xshift=-1.0mm,yshift=0.0mm}:$-$}} }
\tikzset{sumpp/.style={sum,
				label={{above left,xshift=0.0mm,yshift=-1.0mm}:$+$},
				label={{below right,xshift=-1.0mm,yshift=0.0mm}:$+$}} }
\tikzset{sumpn/.style={sum,
				label={{below left,xshift=0.0mm,yshift=1.0mm}:$+$},
				label={{above right,xshift=-1.0mm,yshift=0.0mm}:$-$}} }
\tikzset{sumxpxp/.style={sum,
				label={{above left,xshift=0.0mm,yshift=-1.0mm}:$+$},
				label={{above right,xshift=-0.0mm,yshift=-1.0mm}:$+$}} }
\tikzset{gain/.style={rectangle,draw,very thick,inner sep=2mm,shadow}}

@@

% {{{ Regulator Control System
\clearpage
\section{Regulator Control System}

What is needed is a regulator system where the input is applied
as torque.
The general system is shown in Figure \ref{fig:rpm1}.

% {{{ fig:rpm1, Regulator System
\begin{figure}[hbp!]
\begin{center}

\begin{tikzpicture}[>=triangle 60, scale=0.8,
	node distance=14mm, auto]

	\node[gain] 	(engine)	[]					{engine};
	\node[gain] 	(control)	[below=of engine]	{controller};

	\draw [<-] (engine.north) -- ++(0,10mm) node[above]{$\text{torque}$};
	\draw [->] (engine.east) node[above,xshift=7mm]{rpm}
					-- ++(20mm,0)
					|- (control.east);
	\draw [->] (control.west)
					-- ++(-20mm,0)
					|- (engine.west) node[above,xshift=-9mm]{control};
\end{tikzpicture}

\end{center}
\caption{General system for regulating engine rpm.}\label{fig:rpm1}
\end{figure}
% }}}

@@

@@

\begin{align}
y &= 1450.1 x \label{eq:stable}
\end{align}

>>>>

\usepackage{tikz}
\usetikzlibrary{calc,arrows,positioning,shadows}
%
% control system tikz objects
\tikzset{shadow/.style={drop shadow,fill=white}}
\tikzset{sum/.style={circle,draw,very thick,shadow,minimum size=5mm}}
\tikzset{sumnp/.style={sum,
				label={{above left,xshift=0.0mm,yshift=-1.0mm}:$+$},
				label={{below right,xshift=-1.0mm,yshift=0.0mm}:$-$}} }
\tikzset{sumpp/.style={sum,
				label={{above left,xshift=0.0mm,yshift=-1.0mm}:$+$},
				label={{below right,xshift=-1.0mm,yshift=0.0mm}:$+$}} }
\tikzset{sumpn/.style={sum,
				label={{below left,xshift=0.0mm,yshift=1.0mm}:$+$},
				label={{above right,xshift=-1.0mm,yshift=0.0mm}:$-$}} }
\tikzset{sumxpxp/.style={sum,
				label={{above left,xshift=0.0mm,yshift=-1.0mm}:$+$},
				label={{above right,xshift=-0.0mm,yshift=-1.0mm}:$+$}} }
\tikzset{sumxpxn/.style={sum,
				label={{above left,xshift=0.0mm,yshift=-1.0mm}:$+$},
				label={{above right,xshift=-0.0mm,yshift=-1.0mm}:$-$}} }
\tikzset{gain/.style={rectangle,draw,very thick,inner sep=2mm,shadow}}

@@

% {{{ Control Law, Intuitive Design
\clearpage
\section{Control Law, Intuitive Design}

% {{{ fig:clid01
\begin{figure}[hpb!]
\begin{center}

	\node[gain]  	(Gz)	[] 	{$G(z)$};
	\node[gain] 	(K)	[below=of Gz] 	{$K$};

	 \draw [->]
	 		(Gz.east) -- ++(15mm,0)
			-| ++(0,-10mm)
			|- (K.east);

	 \draw [<-]
	 		(Gz.west) -- ++(-15mm,0)
			-| ++(0,-10mm)
			|- (K.west);
\end{tikzpicture}

\caption{Control Law regulator system.
In typical designs a negative feedback is produced using a negative $K$.}
\label{fig:clid01}
\end{figure}
% }}}

A Control Law Design\autocite[Pg. 280]{franklin1998digital} configured
as a regulator is shown in Figure \ref{fig:clid01}.
Notice that there is no input causing it regulate at a constant set point.
However there are several requirements that must be met in order
for this system to be stable.
First the input to $G(z)$ must reflect the output without any offset.
And second the system will always stabilize at zero.

These requirements can be verified intuitively.
Suppose for an input of $1$ to $G(z)$ and output of $5$ is produced.
What value of $K$ will produce a stable value?

\begin{align*}
	5 K &= 1 \\
	K &= \frac{1}{5}
\end{align*}

Now suppose a disturbance is introduced causing the output to increase
to $6$, will the system stabilize?

\begin{align*}
	6 \cdot K &= y \\
	6 \cdot \frac{1}{5} &= \frac{6}{5} \\
	\frac{6}{5} &> 1 \quad \text{(diverge)}
\end{align*}

An increase in the output creates positive feedback which increases
the input and causes the system diverge.
This exercise can be repeated with a controller that has no offset at zero.
The result is a negative feedback system that will stabilize at zero.

The linear engine model with zero torque (Equation \ref{eq:lems})
supports none of these requirements for a stable Control Law regulator.
The input does not reflect the output; the input range is from zero to
one whereas the output range is from 200 to 1200.
And operation at zero rpm is not realistic.

% {{{ fig:clid02
\begin{figure}[hpb!]
\begin{center}

	\node[sumxpxp]  	(sum1)	[]				{};
	\node[gain]  	(R)		[right=of sum1] 	{$R$};
	\node[gain]  	(Gz)	[right=of R] 		{$G(z)$};
	\node[sumxpxn]	(sum2)	[right=of Gz]		{};
	\node[gain] 	(K)		[below=of Gz] 		{$-K$};

	\draw [<-] (sum1.north) -- ++(0,10mm) node[above]{$\text{rpm}$};
	\draw [->] (sum1.east) -- (R.west);
	\draw [->] (R.east) -- (Gz.west);
	\draw [->] (Gz.east) -- (sum2.west);
	\draw [<-] (sum2.north) -- ++(0,10mm) node[above]{$\text{rpm}$};
	\draw [->] (sum2.east) -- ++(10mm,0) node[above,xshift=-3mm]{$y$}
		-| ++(0,-10mm)
		|- (K.east);
	 \draw [<-] 
	 		(sum1.west) -- ++(-10mm,0) node[above,xshift=3mm]{$u$}
			|- (K.west);
\end{tikzpicture}

\caption{Compensations added to the linear engine model with zero
torque ($G(z)$) to construct a regulator system.
The transfer function $R$ converts rpm input to control output (0 - 1).
The rpm input is the desired idle speed such as 700.
Notice that both the output ($y$) and input ($u$) will be zero when
the desired rpm is reached.}
\label{fig:clid02}
\end{figure}
% }}}

The system in Figure \ref{fig:clid02} adds compensations to the
engine model so that a regulator system can be constructed.
Figure \ref{fig:cl_plot1} shows the response of this system.
While the response isn't very good the design was trivial
by simply choosing an arbitrary constant of $0.01$ for $K$.
Later designs will improve performance by elaborating upon this
system and developing more complex values for $K$.

% {{{ fig:dd_plot2
\begin{figure}[htbp!]
\begin{center}
\includegraphics[scale=0.6]{../octave/cl_ff01}
\end{center}
\caption{Output response of Control Law regulator with $K = 0.01$
and rpm = $700$.
Rpm was allowed to stabilize at 200 rpm before applying a step
to 700 rpm.
Source code is provided in Appendix \ref{app:cl_matlab}.}
\label{fig:cl_plot1}
\end{figure}
% }}}

%Instead of a servo system, which accepts a control input,
%a regulator system can be constructed which tries to maintain
%a single output value.  The general design is shown
%in Figure \ref{fig:reg1}.

% {{{ fig:reg1, Regulator System
%\begin{figure}[hbp!]
%\begin{center}
%
%\begin{tikzpicture}[>=triangle 60, scale=0.8,
%	node distance=14mm, auto]
%
%	\node[gain] 	(engine)	[]					{engine};
%	\node[gain] 	(control)	[below=of engine]	{controller};
%
%	\draw [<-] (engine.north) -- ++(0,10mm) node[above]{$\text{torque}$};
%	\draw [->] (engine.east) node[above,xshift=7mm]{rpm}
%					-- ++(20mm,0)
%					|- (control.east);
%	\draw [->] (control.west)
%					-- ++(-20mm,0)
%					|- (engine.west) node[above,xshift=-9mm]{control};
%\end{tikzpicture}
%
%\end{center}
%\caption{Regulator system for idle rpm control.}\label{fig:reg1}
%\end{figure}
% }}}

%One way to build this regulator system is to use the
%Control Law Design\autocite[Pg. 280]{franklin1998digital}.
%The block diagram of the system is shown in Figure \ref{fig:clff1}.

% {{{ fig:clff1
%\begin{figure}[hpb!]
%\begin{center}
%
%\begin{tikzpicture}[>=triangle 60,
%	node distance=14mm, auto]
%
%	\node[gain]  	(Gz)	[] 	{$\begin{matrix}
%									\dot{x} = Ax + Bu \\
%										 y = Cx + Du
%								 \end{matrix}$};
%	\node[gain] 	(K)	[below=of Gz] 	{$-K$};
%
%	 \draw [->]
%	 		(Gz.east) -- ++(15mm,0)
%			-| ++(0,-10mm)
%			|- (K.east);
%
%	 \draw [<-]
%	 		(Gz.west) -- ++(-15mm,0)
%			-| ++(0,-10mm)
%			|- (K.west);
%
%\end{tikzpicture}
%
%\end{center}
%
%\caption{Control Law system with full feedback.
%It behaves as a regulator since not input is available.}
%\label{fig:clff1}
%\end{figure}
% }}}

@@

\begin{align}
y &= 1450.1 x \label{eq:stable} \\
	y &: \text{rpm (200 - 1200)} \notag \\
	x &: \text{control (0 - 1)} \notag
\end{align}

@@

% {{{ Control Law Matlab Source
\clearpage
\section{Control Law Matlab Source}
\label{app:cl_matlab}

Matlab source code used to the Control Law regulator.

\lstinputlisting[
	caption={Matlab source used to plot the Control Law regulator.}
	]{../octave/cl_ff01.m}
